High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2021-08-02 paper/main.tex
final edits for paper

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data}
  \caption{Recovering known dynamic first-order correlations
      from synthetic data.  Each panel displays the average
    correlations between the vectorized upper triangles of the
    recovered correlation matrix at each timepoint and either the true
    underlying correlation at each timepoint or a reference
    correlation matrix.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets). Different colors denote different kernel shapes, and
    the shading within each color family denotes the kernel width
    parameter.  For a complete description of each synthetic dataset,
    see Synthetic data: simulating dynamic first-order
      correlations.  a. Constant correlations.  These
    datasets have a stable (unchanging) underlying correlation matrix.
    b. Random correlations.  These datasets are generated
    using a new independently drawn correlation matrix at each new
    timepoint. c. Ramping correlations.  These datasets are
    generated by smoothly varying the underlying correlations between
    the randomly drawn correlation matrices at the first and last
    timepoints.  The left panel displays the correlations between the
    recovered dynamic correlations and the underlying ground truth
    correlations.  The middle panel compares the recovered
    correlations with the first timepoint's correlation
    matrix.  The right panel compares the recovered correlations with
    the last timepoint's correlation matrix.
    d. Event-based correlations.  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.  The vertical lines
    denote event boundaries. Source data are provided as a Source Data file.}
  \label{fig:synthetic}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{Recovery of simulated first-order and second-order dynamic correlations.  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices at each
    timepoint and for each synthetic dataset.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see Synthetic data: simulating dynamic higher-order
      correlations.  All estimates represented in this figure were
    computed using a Laplace kernel
    (width~$=20$).  Constant.  These datasets
    have stable (unchanging) underlying second-order correlation
    matrices.  Random. These datasets are
    generated using a new independently drawn second-order correlation
    matrix at each timepoint. Ramping.
    These datasets are generated by smoothly varying the underlying
    second-order correlations between the randomly drawn correlation
    matrices at the first and last timepoints. Event.  These datasets are each generated using five
    randomly drawn second-order correlation matrices that each remain
    stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.  Note that the ``dips''
    and ``ramps'' at the boundaries of sharp transitions (e.g., the
    beginning and ends of the ``constant'' and ``ramping'' datasets,
    and at the event boundaries of the ``event'' datasets) are
    finite-sample effects that reflect the reduced numbers of samples
    that may be used to accurately estimate correlations at sharp
    boundaries.}
  \label{fig:higher_order_sims}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.
    a.~Decoding accuracy as a function of order: PCA.
    Order ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    Feature weighting and testing).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see Identifying robust decoding
      results).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$; these chance accuracies
    were subtracted from the observed accuracies to obtain the
    relative accuracies reported on the $y$-axis).  The error ribbons
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upward
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed $t$-test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space. c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality. This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
   d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality. This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space. See Figures~\kernelshape~and~\kernelwidth~for decoding results broken down by kernel shape and width, respectively.}
  \label{fig:decoding}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{Statistical summary of decoding accuracies for
    different neural features.  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature in
  a decoder that included all feature types (see Feature
    weighting and testing).}
\label{fig:ttests}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/most_abs}
  \caption{Top terms associated with the most strongly
      correlated nodes at each order.  Each color corresponds to one order of
    inter-subject functional correlations.  To calculate the dynamic
    correlations, eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space at each previous order, which allows us
    the map the brain regions at each order by retaining the
    features of the original space. The inflated brain plots
    display the locations of the endpoints of the 10 strongest
    (absolute value) correlations at each order, thresholded at 0.999,
    and projected onto the cortical surface~\citep{CombEtal19}.  The
    lists of terms on the right display the top five Neurosynth
    terms~\citep{RubiEtal17} decoded from the corresponding brain maps
    for each order.  Each row displays data from a different
    experimental condition.  Additional maps and their corresponding
    Neurosynth terms may be found in the Supplementary
      materials (intact: Fig.~\intact; paragraph: Fig.~\para; word:
    Fig.~\word; rest: Fig.~\rest).}
  \label{fig:neurosynth}
\end{figure}

>>>>

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data}
  \caption{Recovering known dynamic first-order correlations
      from synthetic data.  Each panel displays the average
    correlations between the vectorized upper triangles of the
    recovered correlation matrix at each timepoint and either the true
    underlying correlation at each timepoint or a reference
    correlation matrix.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals of the mean (taken across
    datasets). Different colors denote different kernel shapes, and
    the shading within each color family denotes the kernel width
    parameter.  For a complete description of each synthetic dataset,
    see Synthetic data: simulating dynamic first-order
      correlations.  a. Constant correlations.  These
    datasets have a stable (unchanging) underlying correlation matrix.
    b. Random correlations.  These datasets are generated
    using a new independently drawn correlation matrix at each new
    timepoint. c. Ramping correlations.  These datasets are
    generated by smoothly varying the underlying correlations between
    the randomly drawn correlation matrices at the first and last
    timepoints.  The left panel displays the correlations between the
    recovered dynamic correlations and the underlying ground truth
    correlations.  The middle panel compares the recovered
    correlations with the first timepoint's correlation
    matrix.  The right panel compares the recovered correlations with
    the last timepoint's correlation matrix.
    d. Event-based correlations.  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.  The vertical lines
    denote event boundaries. Source data are provided as a Source Data file.}
  \label{fig:synthetic}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{Recovery of simulated first-order and second-order dynamic correlations.  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices at each
    timepoint and for each synthetic dataset.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals of the mean (taken across
    datasets).  For a complete description of each synthetic dataset,
    see Synthetic data: simulating dynamic higher-order
      correlations.  All estimates represented in this figure were
    computed using a Laplace kernel
    (width~$=20$).  Constant.  These datasets
    have stable (unchanging) underlying second-order correlation
    matrices.  Random. These datasets are
    generated using a new independently drawn second-order correlation
    matrix at each timepoint. Ramping.
    These datasets are generated by smoothly varying the underlying
    second-order correlations between the randomly drawn correlation
    matrices at the first and last timepoints. Event.  These datasets are each generated using five
    randomly drawn second-order correlation matrices that each remain
    stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.  Note that the ``dips''
    and ``ramps'' at the boundaries of sharp transitions (e.g., the
    beginning and ends of the ``constant'' and ``ramping'' datasets,
    and at the event boundaries of the ``event'' datasets) are
    finite-sample effects that reflect the reduced numbers of samples
    that may be used to accurately estimate correlations at sharp
    boundaries. Source data are provided as a Source Data file.}
  \label{fig:higher_order_sims}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.
    a.~Decoding accuracy as a function of order: PCA.
    Order ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    Feature weighting and testing).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see Identifying robust decoding
      results).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$; these chance accuracies
    were subtracted from the observed accuracies to obtain the
    relative accuracies reported on the $y$-axis).  The error ribbons
    denote 95\% confidence intervals of the means across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upward
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed $t$-test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals of the mean).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space. c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality. This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
   d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality. This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space. See
    Figures~\kernelshape~and~\kernelwidth~for decoding results broken
    down by kernel shape and width, respectively. Source data are provided as a Source Data file.}
  \label{fig:decoding}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{Statistical summary of decoding accuracies for
    different neural features.  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature in
  a decoder that included all feature types (see Feature
    weighting and testing). Source data are provided as a Source Data file.}
\label{fig:ttests}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/most_abs}
  \caption{Top terms associated with the most strongly
      correlated nodes at each order.  Each color corresponds to one order of
    inter-subject functional correlations.  To calculate the dynamic
    correlations, eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space at each previous order, which allows us
    the map the brain regions at each order by retaining the
    features of the original space. The inflated brain plots
    display the locations of the endpoints of the 10 strongest
    (absolute value) correlations at each order, thresholded at 0.999,
    and projected onto the cortical surface~\citep{CombEtal19}.  The
    lists of terms on the right display the top five Neurosynth
    terms~\citep{RubiEtal17} decoded from the corresponding brain maps
    for each order.  Each row displays data from a different
    experimental condition.  Additional maps and their corresponding
    Neurosynth terms may be found in the Supplementary
      materials (intact: Fig.~\intact; paragraph: Fig.~\para; word:
    Fig.~\word; rest: Fig.~\rest). Source data are provided as a Source Data file.}
  \label{fig:neurosynth}
\end{figure}
